---
title: "Group 4 Assignment 3"
author: "Ajay Arora, Romerl Elizes, Jimmy Ng, Joshua Registe, Adam Rich"
date: "April 4, 2021"
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---



```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE)

packages <- c(
  'tidyverse', 
  'corrplot', 
  'palmerpenguins',
  'class',
  'kableExtra',
  'naniar',
  'DataExplorer',
  'caret',
  'tidymodels',
  'rsample',
  'themis'
  # 'htmlTable', 
  # 'gmodels', 
  # 'car', 
  # 'mice', 
  # 'randomForest', 
  # 'tidyselect', 
  # 'skimr', 
  # 'tidymodels', 
  # 'broom', 
  # 'dotwhisker', 
  # 'vip', 
  # 'parsnip', 
  # 'workflows', 
  # 'recipes', 
  # 'tune', 
  # 'yardstick'
)

for (pkg in packages) {
  suppressPackageStartupMessages(suppressWarnings(
    library(
      pkg, character.only = TRUE, 
      warn.conflicts = FALSE, quietly = TRUE)
  ))
}

# A ggplot2 function
defaulttheme <- theme(
  panel.background = element_blank(),
  panel.border = element_rect(color = "black", fill = NA))

```





## (Q1) k-Nearest Neighbors {.tabset .tabset-fade .tabset-pills}


> [Using] the Penguin dataset . . . please use [the] K-nearest neighbors algorithm to predict the species variable.

The **KNN Algorithm** is a simple classification algorithm.
It stores the training data and classifies
new cases by a majority vote of the case's k-nearest neighbors.

The algorithm is sensitive to the units of the chosen predictor variables.
It is important that we transform all the values to a common scale.
A frequently used practice is to normalize the data to the interval [0, 1].

KNN works well when all predictor data has been converted to 
numerics with the exception of the target variable `species`.

*A note on data preparation:* this is the third time we've worked
with the `palmerpenguins` data.  We've repeated some data preparation
and EDA in the appendix.  The data prep we do here are items 
**specific to the k-NN algorithm**.




```{r}
penguins_na_omit <- readr::read_rds('data/penguins_na_omit.Rds')

normalize <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

penguins_df_norm <- penguins_na_omit %>% 
  select(-year) %>% 
  mutate_if(is.factor, as.integer) %>% 
  mutate_if(is.numeric, normalize) %>% 
  mutate(species = penguins_na_omit$species)

```






### Create Training and Test sets

The first is used to train the system, while the second is used to evaluate the learned or trained system.
In practice, the division of your data set into a test and a training sets is disjoint:
the most common splitting choice is to take 2/3 of your original data set as the training set, while the 1/3 that remains will compose the test set.


```{r}
# Create sample data for testing and training.
set.seed(1234)
ind <- sample(2, nrow(penguins_df_norm), replace=TRUE, prob=c(0.67, 0.33))

# Training and test sets
penguin.train <- penguins_df_norm %>% filter(ind == 1) %>% select(-species)
penguin.test  <- penguins_df_norm %>% filter(ind == 2) %>% select(-species)

# Training and test labels
penguin.trainLabels <- penguins_df_norm$species[ind == 1]
penguin.testLabels <- penguins_df_norm$species[ind == 2]
```





### Build KNN Classifier

This step performs the actual classification.
Note: the k parameter is often an odd number to avoid ties in the voting scores.
A good starting point for k is the square root of the total number of observations.

```{r}
penguin_pred <- knn(
  k = 19,
  train = penguin.train, 
  test = penguin.test, 
  cl = penguin.trainLabels)

caret::confusionMatrix(
  penguin_pred, 
  penguin.testLabels)
```

With k = 19 we have zero misclassifications!



What are the accuracy measures for other values of k?

```{r}
k_search <- function(k) {
  penguin_pred <- knn(
    k = k,
    train = penguin.train, 
    test = penguin.test, 
    cl = penguin.trainLabels)

  cm <- caret::confusionMatrix(
    penguin_pred, 
    penguin.testLabels)
  
  data.frame(
    k = k, 
    Adelie = cm$byClass[1, 11],
    Chinstrap = cm$byClass[2, 11],
    Gentoo = cm$byClass[3, 11] )
}

res <- NULL
for (i in seq(1, 21, 2)) {
  res_i <- k_search(i)
  if (is.null(res))
    res <- res_i
  else
    res <- rbind(res, res_i)
}

print(res)
```

We see that k = 17 is the first value of k for which we have perfect accuracy.
But, the accuracy is extremely good for all values of k.







### What if we do not know `island`?

From previous homework assignments, we found that `island`
was an extremely helpful variable because not all species
are found on each island.  If we wanted to build a model
without `island` (for individuals we might find at sea, for example),
how well would it perform?

```{r}
cols <- c(
  "bill_length_mm", "bill_depth_mm", 
  "flipper_length_mm", "body_mass_g", "sex")

k_search <- function(k) {
  penguin_pred <- knn(
    k = k,
    train = penguin.train[, cols], 
    test = penguin.test[, cols], 
    cl = penguin.trainLabels)

  cm <- caret::confusionMatrix(
    penguin_pred, 
    penguin.testLabels)
  
  data.frame(
    k = k, 
    Adelie = cm$byClass[1, 11],
    Chinstrap = cm$byClass[2, 11],
    Gentoo = cm$byClass[3, 11] )
}

res <- NULL
for (i in seq(1, 21, 2)) {
  res_i <- k_search(i)
  if (is.null(res))
    res <- res_i
  else
    res <- rbind(res, res_i)
}

print(res)
```

With k as low as 3 we get a pretty good model without using `island`.
However, we do not see perfect classification as we do when
the variable is included.





## Appendix: `palmerpenguins` data prep {.tabset .tabset-fade .tabset-pills}


```{r}
library(tidyverse)
library(corrplot)
library(palmerpenguins)
penguins <- palmerpenguins::penguins
```

This data has 344 rows of 8 variables:

| variable           | definition                                                          |
|--------------------|---------------------------------------------------------------------|
| species            | a factor denoting penguin species (Adélie, Chinstrap and Gentoo)    |
| island             | a factor denoting island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)    |
| bill_length_mm     | a number denoting bill length (millimeters)                         |
| bill_depth_mm      | a number denoting bill depth (millimeters)                          |
| flipper_length_mm  | an integer denoting flipper length (millimeters)                    |
| body_mass_g        | an integer denoting body mass (grams)                               |
| sex                | a factor denoting penguin sex (female, male)                        |
| year               | an integer denoting the study year (2007, 2008, or 2009)            |


```{r}
summary(penguins)
```

While there are several methods for dealing with missing data,
we have opted to simply remove rows with NAs.

* imputing numerical variables with `mean` would decrease variance 
* four variables `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g` have the same number of NAs
* removing rows with NAs treates this variables equally
* variable `sex` has 11 missing data points
* we removed these observations because our data is split evenly between male and female


```{r}
penguins_na_omit = na.omit(penguins)
```

There is some correlation in the numeric variables,
as we have seen in our other assignments.

```{r fig.height = 9, fig.width = 12, fig.align = "center"}
penguins_numeric <- penguins_na_omit %>% 
  select(-species, -sex, -island, -year)

# Cor-relation between numerics
corrplot(cor(penguins_numeric), type = 'lower', diag = FALSE)
```

Year is not a potential predictor for this problem so we have excluded it.
There is a strong relationship between `body_mass_g` and `flipper_length_mm`.
The pairs of measurement variables that have absolute correlations above 0.5
we list here:

```{r}
corr_matrtix <- cor(penguins_numeric)

corr_matrtix %>% 
  as.data.frame %>% 
  mutate(x = rownames(corr_matrtix)) %>% 
  pivot_longer(
    cols = -x,
    names_to = 'y', 
    values_to = 'corr') %>% 
  filter(x < y) %>% 
  filter(corr > 0.5 | corr < -0.5) %>% 
  arrange(desc(abs(corr)))
```

We save the prepared data for use in Question #1.


```{r echo=TRUE, eval=FALSE}
readr::write_rds(penguins_na_omit, 'data/penguins_na_omit.Rds')
```







